---
title: "Brute Force KNN parallelisation experiment"
author: "Marc van Heerden (1769865)"
date: "3/8/2018"
output:
    pdf_document:
        latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Overview

* A general introduction to the problem, algorithm, and its various applications. Cite the sources properly
when applicable.
* The solution to the problem, i.e., methodology.
* Experiment setup – the description of data being used; the specifications of machine you run your tests
on; the details of the tests you run; and most importantly, the results. Note: Do not just put your results
in a table or a graph without giving any description of those in the text.
* A summary of your work, plus some discussions on the challenges you encountered in the current
solutions and other potential problems.

## Parallelisation

### Foster's methodology


### Results
More processes, more fun ...

```{r cars}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(knitr))

exp_data <- suppressMessages(readr::read_csv("exp20180308.csv"))

median_data <- 
    exp_data %>%
    mutate(time = time_dist + time_sort) %>%
    group_by(m,n, distance_function, sort_function, serial_parallel) %>%
    summarise(time = median(time),
              time_dist = median(time_dist),
              time_sort = median(time_sort)) %>%
    ungroup
```


```{r}

median_data %>%
    filter(sort_function == "quick",
           distance_function == "euclidean",
           m == 450L) %>%
    kable(caption = "results boyeeee")

```

```{r}

median_data %>%
    filter(sort_function == "quick",
           distance_function == "euclidean") %>%
    ggplot(aes(n, time, color = serial_parallel)) +
        geom_point() +
        facet_grid(m~.) +
        scale_y_log10(name = "time (log10 scale)", 
                      breaks = c(0.03, 0.1, 0.3, 1, 3))

```

## Sorting algorithms


## Distance functions



## System specifications

```{r}
Sys.info()[1:3]
system("lscpu", intern=TRUE)[1:13]
```

